A Fredholm operator approach to the plate buckling eigenvalue problem

### 2018-06-14 doc/draft-01/biharm-evals.tex
Updating tex file

\def\br{{\boldsymbol r}}
\def\bn{{\boldsymbol n}}
\def\xx{{\boldsymbol x}}
\def\bx{{\boldsymbol x}}
\def\yy{{\boldsymbol y}}
\def\by{{\boldsymbol y}}
\def\ss{{\boldsymbol s}}
\def\tt{{\boldsymbol t}}
\def\uu{{\boldsymbol u}}
\def\bu{{\boldsymbol u}}
\def\bl{{\boldsymbol \ell}}
\def\ff{{\boldsymbol f}}
\def\GG{{\bf G}}
\def\Gbh{{G_{BH}}}
\def\II{{\bf I}}
\def\TT{{\bf T}}
\def\SS{{\bf S}}
\def\DD{{\bf D}}
\def\bsigma{{\boldsymbol \sigma}}
\def\bmu{{\boldsymbol \mu}}
\def\bnu{{\boldsymbol \nu}}
\def\btau{{\boldsymbol \tau}}
\def\bigo{\mathcal{O}}
\def\littleo{o}

\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}

@@

Integral equation methods.
\begin{itemize}
\item Book on elastodynamic eigenvalue problems
  \cite{kitahara2014boundary}
\end{itemize}
\section{Mathematical Preliminaries}

\subsection{Notation}

Bold symbols are vectors or tensors.
$\xx = (x_1,x_2)^\intercal$ then $\xx^\bot = (-x_2,x_1)^\intercal$.
Similarly, $\nabla^\bot = (-\partial_{x_2},\partial_{x_1})^\intercal$

\subsection{Green's functions}

$\Delta$
\begin{equation}
  G_L(\xx,\yy) = -\frac{1}{2\pi} \log |\xx-\yy|
\end{equation}

$\Delta + k^2$
\begin{equation}
  G_H(\xx,\yy) = \frac{i}{4} H_0^{(1)}(k|\xx-\yy|)
\end{equation}

$\Delta(\Delta + k^2)$
\begin{equation}
  G_{BH}(\xx,\yy) = -\frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right )
\end{equation}

\section{Integral Equation Formulation}

\subsection{The Helmholtz Stokeslet formulation}

@@

As mentioned in the introduction, there are existing 
integral representations for the solution of the Dirichlet problem for the Helmholtz-biharmonic equation
\begin{align}
 \Delta (\Delta + k^2) w &= 0 \mbox{ in } D \; , \label{eq:biharmD1}\\
 w &= f \mbox{ on }
 \Gamma \label{eq:biharmD2} \; , \\ 
 \frac{\partial w}{\partial n} &= g \mbox{ on } 
 \Gamma \, , \label{eq:biharmD3}
\end{align}
In \cite{Farkas89}, 
the solution is given as the sum of two layer potentials, i.e.
\begin{equation}
  w(\xx) = \int_\Gamma K_1^F (\xx,\yy) \sigma_1(\yy) \, dS(\yy)
  + \int_\Gamma K_2^F(\xx,\yy) \sigma_2(\yy) \, dS(\yy) \; ,
  \label{eq:farkRep}
\end{equation}
where $\sigma_1$ and $\sigma_2$ are unknown densities.

The integral kernels, $K_1^F$ and $K_2^F$ are based on
derivatives of the Green's function for the biharmonic equation.
For two points on the plane, $\xx$ and $\yy$, let $\br = \yy - \xx$. Then 
\begin{align}
  K_1^F(\xx,\yy) &= \partial_{n_{y}n_{y}n_{y}}\Gbh(\xx,\yy)
  + 3 \partial_{n_{y}\tau_{y}\tau_{y}}\Gbh(\xx,\yy) + k^{2} \partial_{n_{y}}\Gbh(\xx,\yy)\; , \label{eq:kf1} \\
  K_2^F(\xx,\yy) &= -\partial_{n_{y}n_{y}}\Gbh(\xx,\yy)
  + \partial_{\tau_{y}\tau_{y}}\Gbh(\xx,\yy) \label{eq:kf2} \; .
\end{align}
More explicitly, we have

\begin{align}
  K_1^F(\xx,\yy) &=  -\dfrac{\br \cdot \bn(\yy)}{|\br|^2}  \cdot \left( \frac{3 i}{2} H_0^{(1)}(k|\br|) - \frac{5 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{6}{\pi k^2 |\br|^2} + \frac{1}{2\pi}  \right) +     \\
  & \quad \dfrac{\left(\br \cdot \bn(\yy)\right)^3}{|\br|^4} \cdot \left( 2i H_0^{(1)}(k|\br|) - \frac{7 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{8}{\pi k^2 |\br|^2}  \right) \, , \nonumber \\
  K_2^F(\xx,\yy) &=  -\left(\frac{1}{2}-\dfrac{\left(\br \cdot \bn(\yy)\right)^2}{|\br|^2} \right)  \cdot \left( \frac{i}{2} H_0^{(1)}(k|\br|) - \frac{i}{k|\br|} H_1^{(1)}(k|\br|) + \frac{2}{\pi k^2 |\br|^2}  \right)\; .
\end{align}

On enforcing the Dirichlet boundary conditions for
$w$, we obtain the integral equation
\begin{equation}
  \begin{split}
    \begin{pmatrix}
      f(\xx) \\
      g(\xx)
    \end{pmatrix} &= 
    \int_\Gamma
    \begin{pmatrix}
      K_{11}^F(\xx,\yy) & K_{12}^F(\xx,\yy) \\
      K_{21}^F(\xx,\yy) & K_{22}^F(\xx,\yy)
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\yy)\\
      \sigma_2(\yy)
    \end{pmatrix}
    \, dS(\yy) \\
    &\quad + \begin{pmatrix}
      1/2 & 0 \\
      -\kappa(\xx) & 1/2
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\xx)\\
      \sigma_2(\xx)
    \end{pmatrix} \; , \label{eq:farkIntEq} 
  \end{split}
\end{equation}
where $\kappa$ denotes the signed curvature as
a function on $\Gamma$ and $\xx$ is a point on $\Gamma$. 
The kernels are given
by $K_{11}^F = K_1^F$, $K_{12}^F = K_2^F$,
\begin{align}
  K_{21}^F(\xx,\yy) &= \left(K_1^F(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{21}^{F}(\xx,\yy) \; \label{eq:farkK21},\\
  K_{22}^F(\xx,\yy) &= \left(K_2^F(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{22}^{F}(\xx,\yy) \label{eq:farkK22} \; .
\end{align}
For a sufficiently smooth and simply connected domain,
the integral equation \cref{eq:farkIntEq} is invertible.
The case of a multiply connected domain is not treated
fully in \cite{Farkas89}, but some of the issues are
considered.

As mentioned above, the kernels $K^F_1$ and $K^F_2$ are
constructed with the goal that the $K^F_{ij}$ are
as smooth as possible. Suppose that the boundary
$\Gamma$ is of class $C^k$. Then, the kernels,
$K^F_{ij}(\xx,\yy)$, are $C^{k-2}$ functions on the boundary
for each $\yy \in \Gamma$.
Therefore, on a smooth
boundary, these kernels are smooth. However, on a
domain with a corner, we note 
that the kernel $K_{21}^F$
has a singularity with strength $r^{-2}$ (see~\cref{fig:ker21plot}). 
This singularity, in addition to the term in \cref{eq:farkIntEq}
which explicitly involves the
curvature, makes the representation \cref{eq:farkRep}
unstable for domains with high curvature (or
corners).

>>>>

\def\br{{\boldsymbol r}}
\def\bh{{\boldsymbol h}}
\def\bn{{\boldsymbol n}}
\def\bT{{\boldsymbol T}}
\def\xx{{\boldsymbol x}}
\def\bx{{\boldsymbol x}}
\def\yy{{\boldsymbol y}}
\def\by{{\boldsymbol y}}
\def\ss{{\boldsymbol s}}
\def\tt{{\boldsymbol t}}
\def\uu{{\boldsymbol u}}
\def\bu{{\boldsymbol u}}
\def\bl{{\boldsymbol \ell}}
\def\ff{{\boldsymbol f}}
\def\GG{{\bf G}}
\def\Gbh{{G_{BH}}}
\def\II{{\bf I}}
\def\TT{{\bf T}}
\def\SS{{\bf S}}
\def\DD{{\bf D}}
\def\bsigma{{\boldsymbol \sigma}}
\def\bmu{{\boldsymbol \mu}}
\def\bnu{{\boldsymbol \nu}}
\def\btau{{\boldsymbol \tau}}
\def\bigo{\mathcal{O}}
\def\littleo{o}
\newcommand\dwdn{\frac{\partial w}{\partial n}}
\newcommand{\cC}{\mathcal C}

\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{lem}{Lemma}

@@

Integral equation methods.
\begin{itemize}
\item Book on elastodynamic eigenvalue problems
  \cite{kitahara2014boundary}
\end{itemize}

Consider the pure bending of an isotropic and homogeneous thin clamped
plate. 
Let $\Omega$ denote the midline of the thin plate, $\Gamma$ denote the 
boundary of $\Omega$ and $w$ denote the vertical displacement.
In this setup, the buckling eigenvalue problem consists of finding 
eigenvalues $k^{2}$ and corresponding non-trivial eigenfunctions
(or buckling modes) $w$ which satisfy the homogeneous Dirichlet
modified Biharmonic problem (with an imaginary parameter $k$) given by
\begin{align}
\Delta (\Delta + k^2) w &= 0 \quad \text{in} \quad \Omega \label{eq:buck1} \\
w &= 0 \quad \text{on} \quad \Gamma \label{eq:buck2}\\
\dwdn &=0 \quad \text{on} \quad \Gamma \label{eq:buck3} \,.
\end{align}
It is well-known from classical potential theory that, 
for the buckling eigenvalue problem, there
exists a countable collection of eigenfrequencies 
$0 < k_{1}^{2} \leq k_{2}^2 \leq \ldots \uparrow \infty$
counting multiplicites.
\begin{remark}
For notational convenience,  
we will refer to the modified Biharmonic problem
with imaginary parameter as the Helmholtz-Biharmonic problem.
\end{remark}

\section{Mathematical Preliminaries}

In this paper, vector-valued
quantities are denoted by bold, lower-case letters
(e.g. $\bh$), while tensor-valued quantities are bold
and upper-case (e.g. $\mathbf{T}$). 
Subscript indices of non-bold characters (e.g. $h_j$ or $T_{jkl}$)
are used to denote the entries within a vector ($\bh$) or tensor ($\bT$).
We use the standard Einstein summation convention; in other words, 
there is an implied sum taken over the repeated indices of 
any term (e.g. the symbol $a_{j} b_{j}$ is used to represent the sum
$\sum_{j} a_{j} b_{j}$).
$\xx = (x_1,x_2)^\intercal$ then $\xx^\bot = (-x_2,x_1)^\intercal$.
Similarly, $\nabla^\bot = (-\partial_{x_2},\partial_{x_1})^\intercal$.

\subsection{The Farkas integral representation}
Following the treatment of \cite{askham}, the fundamental solution
of the Helmholtz-Biharmonic equation is given by
\begin{equation}
  G_{BH}(\xx,\yy) = -\frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right ) \, , 
\end{equation}
where $k$ is the Helmholtz paramater in the Helmholtz-Biharmonic equation,
and $H_{0}^{1}(r)$ is the Hankel function of the first kind of order zero.

In \cite{Farkas89}, Farkas derived integral equation formulations
for the solution of the Dirichlet problem for the biharmonic equation ($k=0$). 
In this formulation, the solution 
$w$ is represented as a sum of two layer potentials
\begin{equation}
  w(\xx) = \int_\Gamma K_1^F (\xx,\yy) \sigma_1(\yy) \, dS(\yy)
  + \int_\Gamma K_2^F(\xx,\yy) \sigma_2(\yy) \, dS(\yy) \; ,
  \label{eq:farkRep}
\end{equation}
where $\sigma_{1}$, and $\sigma_{2}$ are unknown densities, and the kernels
$K_{1}^{F}$, and $K_{2}^{F}$ are linear combinations of derivatives of
the biharmonic Green's function and appropriately chosen to get second kind integral equations
for the unknown densities $\sigma_{1}$ and $\sigma_{2}$ on imposing the boundary conditions.

In \cite{shidongmodbi}, Jiang extended the Farkas formulation for the solution of the modified
biharmonic equation, albeit for different boundary conditions. With minor modifications, their formulation
can also be adopted for the solution of the Dirichlet problem for the Helmholtz-biharmonic equation
\begin{align}
\Delta (\Delta + k^2) w &= 0 \quad \text{in} \quad \Omega \label{eq:helmbi} \\
w &= f \quad \text{on} \quad \Gamma \\
\dwdn &=g \quad \text{on} \quad \Gamma \,,
\end{align}
where $f,g$ are prescribed functions on the boundary.

For two points on the plane, $\xx$ and $\yy$, let $\br = \yy - \xx$. Let the kernels $K_{1}^{F}$, and 
$K_{2}^{F}$ be defined by 
\begin{align}
  K_1^F(\xx,\yy) &= \partial_{n_{y}n_{y}n_{y}}\Gbh(\xx,\yy)
  + 3 \partial_{n_{y}\tau_{y}\tau_{y}}\Gbh(\xx,\yy) + k^{2} \partial_{n_{y}}\Gbh(\xx,\yy)\; , \label{eq:kf1} \\
  K_2^F(\xx,\yy) &= -\partial_{n_{y}n_{y}}\Gbh(\xx,\yy)
  + \partial_{\tau_{y}\tau_{y}}\Gbh(\xx,\yy) \label{eq:kf2} \; .
\end{align}
More explicitly, we have
\begin{align}
  K_1^F(\xx,\yy) &=  -\dfrac{\br \cdot \bn(\yy)}{|\br|^2}  \cdot \left( \frac{3 i}{2} H_0^{(1)}(k|\br|) - \frac{5 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{6}{\pi k^2 |\br|^2} + \frac{1}{2\pi}  \right) +     \\
  & \quad \dfrac{\left(\br \cdot \bn(\yy)\right)^3}{|\br|^4} \cdot \left( 2i H_0^{(1)}(k|\br|) - \frac{7 i}{2 k|\br|} H_1^{(1)}(k|\br|) + \frac{8}{\pi k^2 |\br|^2}  \right) \, , \nonumber \\
  K_2^F(\xx,\yy) &=  -\left(\frac{1}{2}-\dfrac{\left(\br \cdot \bn(\yy)\right)^2}{|\br|^2} \right)  \cdot \left( \frac{i}{2} H_0^{(1)}(k|\br|) - \frac{i}{k|\br|} H_1^{(1)}(k|\br|) + \frac{2}{\pi k^2 |\br|^2}  \right)\; .
\end{align}

On enforcing the Dirichlet boundary conditions for
$w$, we obtain the integral equation
\begin{equation}
  \begin{split}
    \begin{pmatrix}
      f(\xx) \\
      g(\xx)
    \end{pmatrix} &= 
    \int_\Gamma
    \begin{pmatrix}
      K_{11}^F(\xx,\yy) & K_{12}^F(\xx,\yy) \\
      K_{21}^F(\xx,\yy) & K_{22}^F(\xx,\yy)
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\yy)\\
      \sigma_2(\yy)
    \end{pmatrix}
    \, dS(\yy) \\
    &\quad + \begin{pmatrix}
      1/2 & 0 \\
      -\kappa(\xx) & 1/2
    \end{pmatrix}
    \begin{pmatrix}
      \sigma_1(\xx)\\
      \sigma_2(\xx)
    \end{pmatrix} \; , \label{eq:farkIntEq} 
  \end{split}
\end{equation}
where $\kappa$ is the signed curvature at $\xx \in \Gamma$.
The kernels are given
by $K_{11}^F = K_1^F$, $K_{12}^F = K_2^F$,
\begin{align}
  K_{21}^F(\xx,\yy) &= \left(K_1^F(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{21}^{F}(\xx,\yy) \; \label{eq:farkK21},\\
  K_{22}^F(\xx,\yy) &= \left(K_2^F(\xx,\yy)\right )_{n_x} \nonumber \\
  &= \bn(\xx) \cdot \nabla_{\xx} K_{22}^{F}(\xx,\yy) \label{eq:farkK22} \; .
\end{align}

The following lemma shows that the integral equation~\cref{eq:farkIntEq} is invertible for smooth domains
whenever $k$ is not a buckling eigenfrequency.

\begin{lem}
Suppose that $\Omega \subset \R^2$ is a simply connected region whose boundary $\Gamma$ is $\cC^{2}$. 
Furthermore, suppose that $k$ is not a buckling eigenfrequency of $\Omega$, i.e., the only solution $w$ to~\cref{eq:buck1,eq:buck2,eq:buck3} is  $w \equiv 0$. Suppose further that $f,g \in \cC^{2}(\Gamma)$. 
Then there exists a unique solution $\sigma_{1},\sigma_{2} \in \cC^{2}$ which satisfies~\cref{eq:farkIntEq}. 
Furthermore, $w(\bx)$ defined via~\cref{eq:farkRep} satisfies the Helmholtz-Biharmonic equation~\cref{eq:helmbi}
along with the boundary conditions $w(\bx)=f(\bx)$ and $\dwdn(\bx) = g(\bx)$ for $\bx \in\Gamma$. 
\end{lem}

\begin{proof}
The proof is a minor modification of the one presented in~\ref{eq:shidongmodbi} and is contained in~\cref{sec:app}.
\end{proof}

The kernels $K^F_1$ and $K^F_2$ are
constructed with the goal that the $K^F_{ij}$ are
as smooth as possible. Suppose that the boundary
$\Gamma$ is of class $C^k$. Then, the kernels,
$K^F_{ij}(\xx,\yy)$, are $C^{k-2}$ functions on the boundary
for each $\yy \in \Gamma$.
Therefore, on a smooth
boundary, these kernels are smooth. However, on a
domain with a corner, we note 
that the kernel $K_{21}^F$
has a singularity with strength $r^{-2}$ (see~\cref{fig:ker21plot}). 
This singularity, in addition to the term in \cref{eq:farkIntEq}
which explicitly involves the
curvature, makes the representation \cref{eq:farkRep}
unstable for domains with high curvature (or
corners).

\subsection{Helmholtz-Stokes flow in a plane}

@@

As mentioned in the introduction, there are existing 
integral representations for the solution of the Dirichlet problem for the Helmholtz-biharmonic equation
\begin{align}
 \Delta (\Delta + k^2) w &= 0 \mbox{ in } D \; , \label{eq:biharmD1}\\
 w &= f \mbox{ on }
 \Gamma \label{eq:biharmD2} \; , \\ 
 \frac{\partial w}{\partial n} &= g \mbox{ on } 
 \Gamma \, , \label{eq:biharmD3}
\end{align}
In \cite{Farkas89}, 
the solution is given as the sum of two layer potentials, i.e.
where $\sigma_1$ and $\sigma_2$ are unknown densities.

The integral kernels, $K_1^F$ and $K_2^F$ are based on
derivatives of the Green's function for the biharmonic equation.
